---
title: "My project"
author: "Anthony Thieme"
date: "2/10/2021"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## Computational Musicology Portfolio
### what makes a "study playlist" a study playlist
// disclaimer: did not have enough time for this course past week, so all is a bit basic

The music in the "study" category has a really low energy, and a really low valence.


I am thinking that i should maybe research a study playlist vs another, but i am not quite sure how to get going with my project.



```{r libraries}
library(tidyverse)
library(spotifyr)
library(ggridges)
```

## Including Plots

# getting the playlist_tracks

```{r spotify URI}
study_1 <- get_playlist_audio_features("","471N195f5jAVs086lzYglw")

study_lofi <- get_playlist_audio_features("","37i9dQZF1DX8Uebhn9wzrS")

study_lofi_hiphop <- get_playlist_audio_features("","0vvXsWCC9xrXsKd4FyS8kM")

study_intense <- get_playlist_audio_features("","37i9dQZF1DX8NTLI2TtZa6")

study_instumental <- get_playlist_audio_features("","37i9dQZF1DX9sIqqvKsjG8")

study_piano <- get_playlist_audio_features("","37i9dQZF1DX47ov4h4oSjW")




study_relaxed <-
  bind_rows(
    study_1 %>% mutate(category = "Study 1"),
    study_lofi %>% mutate(category = "Chill Lofi Study Beats"),
    study_lofi_hiphop %>% mutate(category = "Lofi hiphop")
    

    )
study_piano <-
  bind_rows(
    study_intense %>% mutate(category = "intense"),
    study_instumental %>% mutate(category = "Instrumental"),
    study_piano %>% mutate(category = "piano")
  )


study_combines <-
  bind_rows(
    study_intense %>% mutate(category = "intense"),
    study_instumental %>% mutate(category = "Instrumental"),
    study_piano %>% mutate(category = "piano"),
    study_1 %>% mutate(category = "Study 1"),
    study_lofi %>% mutate(category = "Chill Lofi Study Beats"),
    study_lofi_hiphop %>% mutate(category = "Lofi hiphop")
)
view(study_combines)
```

``` {r}

study_combines %>%
  ggplot(aes(x= valence, y = energy, color = category)) + geom_point() 
  
  
  
study_combines %>% 
  ggplot(aes(x=valence, y = category, fill = ..x..)) + geom_density_ridges_gradient()

study_combines %>%
  ggplot(aes(x=instrumentalness, y = category, fill = ..x..)) + geom_density_ridges_gradient()

study_combines %>%
  ggplot(aes(x=loudness, y = category, fill = ..x..)) + geom_density_ridges_gradient()

study_combines %>%
  ggplot(aes(x=tempo, y = category, fill = ..x..)) + geom_density_ridges_gradient()
  
  


```
Note that the `echo = FALSE` parameter was added to the code chunk to prevent printing of the R code that generated th