---
title: "Computational Musicology: The 27 Club"
author: "Anthony Thieme"

output: 
  flexdashboard::flex_dashboard:
    storyboard: true
    self_contained: false
    source: https://github.com/toonthieme/ComMus/blob/main/corpus.Rmd
    theme:
      heading_font:
        google: 
          family: Rajdhani
          wght: 700
      base_font:
        google: Fira Sans
      code_font:
        google: Fira Code
      bg: "#FFFFFF"
      fg: "#212529" 
      primary: "#2b2bee"
      secondary: "#39d7b8"
      success: "#39d7b8"
      danger: "#fa5577"
      warning: "#ffb14c"
      info: "#0cc7f1"
---

```{r, echo = FALSE}
knitr::opts_chunk$set(cache = TRUE)
```

```{r, setup}
library(tidyverse)
library(tidymodels)
library(plotly)
library(heatmaply)
library(protoclust)
library(cowplot)
library(spotifyr)
library(compmus)
library(ggplot2)
```

Introduction
=========================================

### Introduction

My corpus focuses on the "27 Club," a collective of influential musicians whose lives tragically ended at the age of 27. This group includes iconic figures such as Jimi Hendrix, Janis Joplin, Jim Morrison, Kurt Cobain, and Amy Winehouse. I chose to analyse Spotify playlists dedicated to these artists to explore a diverse array of musical genres, from rock and blues to grunge and soul. My goal is to delve into their unique styles, lyrical content, and the cultural contexts of their times, aiming to uncover both the stark differences and surprising similarities in their music.

In my project, I intend to reflect on the personal struggles, artistic visions, and societal issues these artists addressed through their music. By focusing on audio features like energy, valence, tempo, danceability, and acousticness extracted from Spotify, I aim to dissect the emotional and energetic dimensions of their tracks. A significant part of my analysis involves examining how musical modes (Major vs. Minor) correlate with the mood and intensity of their music, offering insights into the emotional depth conveyed.

I've highlighted the importance of understanding each artist's stylistic evolution, the thematic continuity across varied genres, and the profound emotional resonance in their music. Including both live performances and studio recordings in my analysis might provide deeper insights into their artistry and the raw emotions expressed in live settings.

This corpus is not just a tribute to the talents and legacies of these artists but also a comprehensive examination of their contributions to music and culture. Through detailed analysis—comparing energy and valence, exploring danceability versus acousticness, and examining the impact of musical mode—I seek to offer a nuanced understanding of the "27 Club" members' enduring influence and the complex tapestry of themes that define their work.

For now i have used the "This is .." playlists for each club member, so there are multiple songs. I still need to decide whether i am going to focus on certain songs of these artists (and which ones those will be).

### into
this also is a text

Visualisation {.storyboard}
=========================================
### Storyboard Vis 1

```{r fetchin_data}

nirvana <- get_playlist_audio_features("","37i9dQZF1DZ06evO3M0Fbi")
hendrix <- get_playlist_audio_features("","37i9dQZF1DZ06evO4cWDcc")
joplin <- get_playlist_audio_features("","37i9dQZF1DZ06evO2Oo4IE")
winehouse <- get_playlist_audio_features("","37i9dQZF1DZ06evO42EWMo")
morrison <- get_playlist_audio_features("","37i9dQZF1DZ06evO19UBIk")

# Combine all tracks into one dataframe and add an artist column
tracks_combined <- bind_rows(
  nirvana %>% mutate(artist = "Nirvana"),
  hendrix %>% mutate(artist = "Jimi Hendrix"),
  joplin %>% mutate(artist = "Janis Joplin"),
  winehouse %>% mutate(artist = "Amy Winehouse"),
  morrison %>% mutate(artist = "Jim Morrison")
)

# Prepare the data by selecting relevant features and renaming appropriately
tracks_prepared <- tracks_combined %>%
  select(track.id, track.name, track.artists, artist, energy, valence, tempo, instrumentalness, danceability, acousticness, liveness, loudness, speechiness, time_signature, mode, key)



```




```{r}
# Energy vs. Valence Plot
ggplot(tracks_prepared, aes(x = valence, y = energy, color = artist)) +
  geom_point(alpha = 0.7) +
  theme_minimal() +
  labs(
    title = "Energy vs. Valence in Music of the '27 Club'",
    x = "Valence (Musical Positiveness)",
    y = "Energy (Intensity and Activity)",
    color = "Artist"
  ) +
  scale_color_brewer(palette = "Set2") +
  theme(
    plot.title = element_text(face = "bold", size = 14),
    axis.title = element_text(size = 12),
    legend.title = element_text(size = 12),
    legend.position = "right"
  )
```

***
**Energy and Valence**: 
This plot compares the energy and valence of tracks by different artists in the "27 Club". It helps in understanding the mood and intensity of their music.



### Storyboard 2

```{r}
# Danceability vs. Acousticness Plot
ggplot(tracks_prepared, aes(x = danceability, y = acousticness, color = artist)) +
  geom_point(alpha = 0.6) +
  theme_minimal() +
  labs(
    title = "Danceability vs. Acousticness in '27 Club' Music",
    x = "Danceability",
    y = "Acousticness",
    color = "Artist"
  ) +
  scale_color_brewer(palette = "Set3") +
  theme(legend.position = "right")

```

***
this is comment about energy and valence

This plot helps to explore how danceable and acoustic the compositions of the "27 Club" artists are, potentially indicating their style's intensity and texture.

Conclusion
==========================
### Conclusion
This will be my conclusion

Column
---------------------------
this is a colimn



